---
title: "climateR"
output: github_document
---

# climateR

<!-- badges: start -->
<!-- badges: end -->

An R package for retrieving monthly climatological data from the Climatempo platform using automated web scraping. The package provides flexible tools to search locations and extract structured climate data for ecological, environmental, and data science applications.

---

## Installation

You can install the development version of **climateR** from GitHub:

```{r, eval=FALSE}
# Install development version from GitHub
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

remotes::install_github("jrz-s/climateR")
```

## Overview

**climateR** enables users to:

- Search for locations using flexible input (region, state, or city)
- Retrieve monthly climatological data (temperature and precipitation)
- Handle partial, case-insensitive, and accent-insensitive queries
- Aggregate data across multiple locations

## Example

```{r, eval=FALSE}
library(climateR)

# Retrieve climate list for Aracaju (Sergipe, Brazil)
df <- search_locations(state = "se", city = "aracaju")
climate_list(df)

# Retrieve climate data for Aracaju (Sergipe, Brazil)
climate_data(state = "se", city = "aracaju")
```

## Input flexibility

The package supports robust matching of user inputs:

- Partial names (e.g., "arac" → "aracaju")
- Case-insensitive matching (e.g., "Sergipe" = "sergipe")
- Accent-insensitive matching (e.g., "São" = "Sao")
- Informative error messages for ambiguous inputs

## Workflow

A typical workflow consists of:

1. Searching locations using `search_locations()`
2. Retrieving climate data using `climate_data()`
3. Performing downstream analysis or visualization

## Output

The main function returns a tidy tibble with the following variables:

`region:` Geographic region
`state:` State abbreviation
`city_name:` City name
`site:` Site identifier (optional)
`month:` Month (1–12)
`tmin:` Minimum temperature (°C)
`tmax:` Maximum temperature (°C)
`precip:` Precipitation (mm)

## Disclaimer

This package performs automated web scraping from the Climatempo website.

Users should:

- Avoid excessive or high-frequency requests
- Respect the terms of service of the data provider
- Use appropriate delays between requests

The authors are not affiliated with Climatempo.

## Author

Developed by Jhonatan Rafael Zárate-Salazar, PhD
Postdoctoral researcher, Graduate Program in Ecology and Conservation
Federal University of Sergipe, São Cristóvão, Sergipe, Brazil

## Collaborators

Sidney Feitosa Gouveia, PhD
Professor, Department of Ecology
Federal University of Sergipe, São Cristóvão, Sergipe, Brazil
